Redshift Evolution of the Underlying Type Ia Supernova Stretch Distribution

### 2020-04-23 Nicolas.tex
Missing commit of Nora changes before MR calll

\abstract{
{The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as survey statistics increase, the question of
astrophysical systematic uncertainties rises,  notably that of
the SN~Ia population redshift drift.}
{In this paper, we study the redshift evolution of SN~Ia
\textsc{\texttt{SALT2.4}} lightcurve stretch, a purely intrinsic SN property, to
probe this drift.}
{The SN stretch has been shown to strongly correlate with the SN's environment, notably with stellar age tracers. Following the evolution of
fraction of young and old SNe~Ia as predicted by \cite{rigault2018}, and
assuming constant underlying stretch
distribution for each population \mr{as a function of redshift}, model the expected SN~Ia stretch
distribution as a function of redshift. We test our prediction against data
from the literature chosen to have negligible selection effects to ensure
that any observed drift is indeed astrophysical and not observational.}
{We clearly demonstrate that the underlying SNe~Ia stretch distribution is
evolving as a function of redshift, and that the young/old drifting model is
a much better description of the data than any non-drifting model, including
the sample-based asymmetric distributions used by the Beams with Bias
Correction algorithm to correct Malmquist bias. Our favored underlying
stretch model is bimodal: a high-stretch mode shared by both young and old
environments, and a low-stretch mode exclusive to old environments.}
\nn{The
precise impact of the redshift evolution of the underlying SNe Ia population on
cosmology remains to be studied}. Yet, the astrophysical drift of the SN stretch
distribution does affect current Malmquist bias corrections and thereby
distances derived from SN affected by selection effects. We highlight that such
a bias will increase with surveys covering larger redshift ranges, which is
particularly important for LSST.}

@@

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
have enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate ($z<0.5$), when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant (H$_0$), provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
the SNe~Ia are anchored on Cepheids \citep[the SH0ES
project,][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
from cosmic microwave background (CMB) data measured by Planck assuming the
standard $\L$CDM \citep{planck2018,riess2019,reid2019}, or when the SN luminosity is anchored
at intermediate redshift by the baryon acoustic oscillation (BAO) scale \citep{feeney2019}. While using the tip of the
red giant branch technique in place of the Cepheids seem to favor an
intermediate value of H$_0$ \citep{freedman2019,freedman2020}, time delay
measurements from strong lensing seem to also favor high H$_0$ values
\citep{wong2019}.

The H$_0$ tension has received a lot of attention, as it could be a sign of new
fundamental physics. Yet, no simple solution is able to accommodate this H$_0$
tension when accounting for all other probes \citep{knox2019}, but see e.g. \cite{poulin2019}.
Alternatively, systematic effects affecting one or several of the aforementioned
analyses could also explain the tension. \cite{rigault2015} suggested that SNe
Ia from the Cepheid calibrator sample differ by construction from the Hubble
flow \nn{sample} ones as \nn{the former} strongly
favors young stellar population\nn{s}, while the latter not. This selection effect would impact the derivation of H$_0$ if
SNe~Ia from young and older environments differ in average standardized
magnitudes.

For the last decades, numerous analyses have studied the relation between SNe~Ia
and host properties, finding first that the standardized SNe~Ia magnitude
significantly depends on the host stellar mass, SNe~Ia from high-mass host being
brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
betoule2014, rigault2018, kim19}.  This mass-step correction is currently used
in cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including
for deriving H$_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
between the SNe and their host remains unclear when using global properties such
as the host stellar mass, which raises the question of the accuracy of the
correction. More recently, studies have used the local SN environment to probe
more direct connections between the SN and its environments \citep{rigault2013},
showing that local age tracers such as the Local specific Star Formation Rate
\nn{(LsSFR)} or the local color are more strongly correlated with the standardized
SN magnitude \citep{rigault2018, roman2018, kim18}, suggesting age as the
driving parameter underlying the mass-step. If true, this would have a
significant impact for cosmology, \mr{since the environmental correction to apply to SN standardization could strongly vary with redshift.} \citep{rigault2013, childress2014, scolnic2018a}. Yet, the
importance of local SN environmental studies remains highly debated
\cite[e.g.][]{jones2015, jones2019} and especially the impact \mr{of such an astrophysical bias} has on the derivation of
H$_0$ \citep{jones2015, riess2016, riess2018, rose2019}.

%\nn{We point that there has been sample-based approach meant to account for an unmodeled variation of host-related parameters thanks to multiple redshift-span-limited surveys \citep{scolnic2016, kessler2017}, that we show aren't as effective as a direct modeling of such a drift}.

The concept of the SNe~Ia age dichotomy arose with the study of the SN~Ia rate. 
\cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have
    shown that the relative SNe~Ia rate in galaxies could be explained if two
    populations existed, one young, following the host star formation activity,
    and one old following the host stellar mass (the so called ``prompt and
    delayed'' or ``A+B'' model). In \cite{rigault2018} we used the LsSFR to classify which
    are the younger (those with a high LsSFR) and which are the older (those
    with low LsSFR). Furthermore, since the first SNe~Ia host analyses, the SN
    stretch has been known to be strongly correlated with the SN host properties
    \citep{hamuy1996, hamuy2000}, correlation that has been extensively confirmed since \citep[e.g.][]{neill2009,
sullivan2010, lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013,
rigault2013, pan2014, kim19}. Following the ``A+B'' model and the connection
between SN stretch and host properties, \cite{howell2007} first discussed the
potential redshift drift of the SN stretch distribution. In this paper we
revisit this analysis with the most recent SNe~Ia dataset.

In this paper, we take a step aside to probe the validity of our modeling of the
SN population, which we claim to be constituted of two age-populations
\citep{rigault2013,rigault2015,rigault2018}: one old and one younger, the former
having on average lower lightcurve stretches and being brighter after
standardization. We use the correlation between the SN age, as probed by the LsSFR, and the SN stretch to model
the expected evolution of the underlying SN stretch distribution as a function
of redshift. This modeling relies on three assumptions: (1) there are two
distinct populations of SNe~Ia; (2) the relative fraction of each of these
populations as a function of redshift follows the model presented in
\cite{rigault2018} and (3) the underlying distribution of stretch for each age
sample is constant. This paper aims at testing this specific model with datasets from the literature. We highlight

We present in section~\ref{sec:sample} the sample we are using for this
analysis, derived from the Pantheon catalog \citep{scolnic2018a}. We discuss the
importance of obtaining a ``complete'' sample, i.e.\ representative of the true
underlying SNe Ia distribution, and how we build one from the Pantheon sample.
We then present in section~\ref{sec:modeling} our modeling of the distribution of stretch and our results are presented in
section~\ref{sec:results}. In this section we test whether the SN stretch distribution evolves as a function of redshift and if the aforementioned age model is in good agreement with this evolution. We briefly discuss these results in the context of SN cosmology in section~\ref{sec:bbc} and we conclude in
section~\ref{sec:ccl}.

@@

We base our analysis on the most recent comprehensive SNe~Ia
compilation, the Pantheon catalog from \cite{scolnic2018a}. A naive approach to
test the SN stretch redshift drift would be to simply compare the observed SN
stretch distributions in a few bins of redshift. In practice, however, \mr{differential} selection effects are affecting the observed SN stretch
distributions. Indeed, because the observed SN~Ia magnitude correlates with the
lightcurve stretch (and color), the first SNe~Ia that a magnitude-limited survey
will miss are the lowest-stretch (and reddest) ones. Consequently, if
magnitude-related selection effects are not accounted for, one might confuse
true population drift with survey properties, and conversely.

@@

We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color of SNe~Ia
from the following surveys: PanStarrs \citep[PS1][]{rest2014}, the Sloan Digital
Sky Survey \citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full
distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011} and
\citet{campbell2013} for similar contours, the second using a more conservative
$|c| \leq 0.2$ cut.  Assuming the SN absolute magnitude with $x_1=0$ and $c=0$
is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$
along the aforementioned ellipse given the standardization coefficient
$\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: the faintest SN~Ia is
that with $(x_1=-1.65, c=0.25)$ and an absolute standardized magnitude at peak
in Bessel $B$ band of $M^{t_0}_{min} = -18.31$~mag. Since one ought to detect
this object typically 5 days before and a week after peak to build a suitable lightcurve, the effective limiting standardized
absolute magnitude is approximately $M_{lim} = -18.00$~mag. Hence, given the
magnitude limit $m_{lim}$ of a magnitude limited survey, one can derive the
maximum redshift $z_{lim}$ above which the faintest SNe~Ia will be missed using
the relation between apparent magnitude, redshift and absolute magnitude
$\mu(z_{lim}) = m_{lim} - M_{lim}$.

% YC: Figs and Tables *after* their 1st ref. in the text!
\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
    \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$)
        lightcurve parameters of SNe~Ia from the SDSS, PS1 and SNLS samples of
        the Pantheon catalog. The individual SNe are shown as blue dots. The ellipse $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by
the corresponding standardized absolute magnitude using the $\alpha$ and $\beta$
coefficients from \cite{scolnic2018a}. The grey diagonal lines represent the
$(x_1, c)$ evolution for $m = m_{lim}$, for $z$ between $0.50$ and $z=1.70$
using SNLS's $m_{lim}$ of $24.8$~mag.}
    \label{fig:maglim}
\end{figure}

@@

In a similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However, the
SDSS surveys were more sensitive to limited spectroscopic resources;
\cite{kessler2009} pointed out that during the first year of SDSS, SNe~Ia with
$r-mag<20.5$ were favored for spectroscopic follow-up, corresponding to a
redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were available, and \cite{kessler2009} and
\cite{dilday2008} show a \mr{reasonable} completeness up to
$z_{lim}=0.2$. Following these analyses, we will use $z_{lim}=0.2$ as the
baseline SDSS redshift limit.

The sample selection is summarized in Table~\ref{tab:sample}, and the redshift distribution of these three surveys is shown in Fig.~\ref{fig:cuts}.
As expected, the selected redshift limits roughly correspond to the peak of
these histograms.

@@

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
and~4.2.2; \mr{SNe~Ia time series are published in \citealt{Saunders2020}, see also, Aldering in prep. and Ponder in prep.)}. The
spectroscopic follow-up was done over a redshift range of $0.02 < z <
0.09$~\citep[as in][]{rigault2018}, while the search was much deeper. As such,
these SNe~Ia are assumed to be a random sampling of the underlying SN
population. The SNfactory sample is particularly useful for studying SN property
drift, as it enables us to have a large SN~Ia sample at $z<0.1$.

@@

We present the stretch distribution and redshift histogram of these five surveys
up to their respective $z_{lim}$ in Fig.~\ref{fig:sample}. \mr{The evolution of the mean stretch is also shown in Fig.~\ref{fig:modelall}, where the data are split in redshift bins of regular sample size. We see that SNe~Ia at higher redshift have on average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying stretch distribution is drifting.}

@@

\subsection{``Base'' underlying stretch distribution}
\label{sec:basemodel}

To model the evolution of the \nn{full} SN stretch distribution as a function of redshift, given our aforementioned model of the evolution of the fraction of
younger and older SNe~Ia with cosmic time, we need to model the SN stretch distribution for each age
subsample.

@@

\mr{The stretch probability distribution function (pdf) of a given SN will be the linear combination of the stretch distributions of these two population weighted by its probability $y^i$ to be young (see Section~\ref{sec:basemodelapplied}). But generally, the fraction of young SNIa as a function of redshift is given by $\delta$(z) (see Eq.~\ref{eq:delta}) and therefore, our redshift drift model of the underlying distribution of SNIa as a function of redshift $X_1(z)$ is given by}:
\begin{multline}\label{eq:stretchz}
    X_1(z) = \delta(z) \times \mathcal{N}(\mu_1,\sigma_1{}^2) + {} \\
    (1-\delta(z)) \times \left[ a\times\mathcal{N}(\mu_1,\sigma_1{}^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2) \right].
\end{multline}

\subsection{Base model applied to data}
\label{sec:basemodelapplied}

Given the probability $y^i$ that a given SN is young, and assuming our Base (see section~\ref{sec:basemodel}), the probability to measure a \textsc{\texttt{SALT2.4}} stretch $x_1^i$ with an error d$x_1^i$ is given by:
\begin{align}\label{eq:likelihoodsnf}
    \prob{x^i_1}{\vec{\theta}; \mathrm{d}x^i_1, y^i} =
    y^i & \times
    \mathcal{N}\left(x^i_1 \mid \mu_1, \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    (1-y^i) &\times \bigg[
    a \times \mathcal{N}\left(x^i_1 \mid \mu_1,
    \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    & (1-a) \times \mathcal{N}\left(x^i_1 \mid \mu_2,
    \sigma_2{}^{2}+\mathrm{d}x^i_1{}^2\right) \bigg],
\end{align}

The maximum-likelihood estimate of the 5~free parameters
$\vec{\theta}\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained
by minimizing the following:
\begin{equation}\label{eq:likelihood}
    -2\ln(L) = -2 \sum_i \ln \prob{x^{i}_{1}}{\vec{\theta};
    \mathrm{d}x^{i}_{1}, y^i}.
\end{equation}

>>>>

\abstract{
{The true nature of type Ia supernovae (SNe~Ia) remains largely unknown, and as
\nn{\sout{measure of}} survey statistic\nn{s} increase, the question of
astrophysical systematic uncertainties rises, \nn{\sout{and}} notably that of
the SN~Ia population redshift drift.}
{In this paper, we study the redshift evolution of SN~Ia
\textsc{\texttt{SALT2.4}} lightcurve stretch, a purely intrinsic SN property, to
probe this drift.}
{The SN stretch has been shown to strongly correlate with the SN's environment,
\nn{\sout{and}} notably with stellar age tracers. Following the evolution of
fraction of young and old SNe~Ia as predicted by \cite{rigault2018}, and
assuming non-drifting \yc{$\to$constant/stationary?} underlying stretch
distribution for each population, model the expected SN~Ia stretch
distribution as a function of redshift. We test our prediction against data
from the literature chosen to have negligible selection effects to ensure
that any observed drift is indeed astrophysical and not observational.}
{We clearly demonstrate that the underlying SNe~Ia stretch distribution is
evolving as a function of redshift, and that the young/old drifting model is
a much better description of the data than any non-drifting model, including
the sample-based asymmetric distributions used by the Beams with Bias
Correction algorithm to correct Malmquist bias. Our favored underlying
stretch model is bimodal: a high-stretch mode shared by both young and old
environments, and a low-stretch mode exclusive to old environments.}
{\sout{The underlying SNe~Ia population is evolving as a function of redshift
    \yc{This is purely a repetition of a previous sentence} \nn{\textit{see new
sentence}} and the exact impact on cosmology remains to be studied.} \nn{The
precise impact of the redshift evolution of the underlying SNe Ia population on
cosmology remains to be studied}. Yet, the astrophysical drift of the SN stretch
distribution does affect current Malmquist bias corrections and thereby
distances derived from SN affected by selection effects. We highlight that such
a bias will increase with surveys covering larger redshift ranges, which is
particularly important for LSST.}
}

@@

\yc{\textit{YC: you should run a grammar corrector, e.g. \texttt{grammarly}.
e.g. "SNe dichotomy" vs. "SN dichotomy", "analysis" vs. "analyses", etc.}}
\nn{\textit{By the way, as we capitalize ``Fig.'' before a reference, shouldn't
we capitalize ``Section'' before reference?}}

Type Ia supernovae (SNe Ia) are powerful cosmological distance indicators that
have enabled the discovery of the acceleration of the Universe's expansion
\citep{riess1998, perlmutter1999}. They remain today a key cosmological probe to
understand the properties of dark energy (DE) as it is the only tool able to
precisely map the recent expansion rate ($z<0.5$), when DE is driving it
\citep[e.g.][]{scolnicastro2020}. They also are key to directly measure the
Hubble Constant (H$_0$), provided one can calibrate their absolute magnitude
\citep{riess2016, freedman2019}. Interestingly, the value of H$_0$ derived when
the SNe~Ia are anchored on Cepheids \citep[the SH0ES
project,][]{riess2009,riess2016} is $\sim5\sigma$ higher than what is predicted
\yc{from} cosmic microwave background (CMB) data measured by Planck assuming the
standard $\L$CDM \yc{ref?} \nn{\textit{isn't it at the end of the sentence? Do
you think it should be before the "or"?}}, or when the SN luminosity is anchored
at intermediate redshift by the baryon acoustic oscillation (BAO) scale
\citep{riess2019,reid2019,planck2018, feeney2019}. While using the tip of the
red giant branch technique in place of the Cepheids seem to favor an
intermediate value of H$_0$ \citep{freedman2019,freedman2020}, time delay
measurements from strong lensing seem to also favor high H$_0$ values
\citep{wong2019}.

The H$_0$ tension has received a lot of attention, as it could be a sign of new
fundamental physics. Yet, no simple solution is able to accommodate this H$_0$
tension when accounting for all other probes \citep{knox2019} and \nn{\sout{the
current most promising}} \nn{one currently possible} scenario appears to be a
burst of expansion at the matter-radiation decoupling moment caused by a (fine
tuned) early dark energy \citep{poulin2019}. \yc{\textit{I don't see the point
of going into this debate...}}

Alternatively, systematic effects affecting one or several of the aforementioned
analyses could also explain the tension. \cite{rigault2015} suggested that SNe
Ia from the Cepheid calibrator sample differ by construction from the Hubble
flow \nn{sample} ones as \nn{\sout{it \yc{they?}}} \nn{the former} strongly
favors young stellar population\nn{s} \nn{\sout{environments where one could
find Cepheids}}. This selection effect would impact the derivation of H$_0$ if
SNe~Ia from young and older environments differ in average standardized
magnitudes.

For the last decades, numerous analyses have studied the relation between SNe~Ia
and host properties, finding first that the standardized SNe~Ia magnitude
significantly depends on the host stellar mass, SNe~Ia from high-mass host being
brighter on average \cite[e.g.][]{kelly2010, sullivan2010, childress2013,
betoule2014, rigault2018, kim19}.  This mass-step correction is currently used
in cosmological analyses \citep[e.g.][]{betoule2014, scolnic2018a}, including
for deriving H$_0$ \citep{riess2016, riess2019}. Yet, the underlying connection
between the SNe and their host remains unclear when using global properties such
as the host stellar mass, which raises the question of the accuracy of the
correction. More recently, studies have used the local SN environment to probe
more direct connections between the SN and its environments \citep{rigault2013},
showing that local age tracers such as the Local specific Star Formation Rate
\nn{(LsSFR)} or the local color are more strongly correlated with the standardized
SN magnitude \citep{rigault2018, roman2018, kim18}, suggesting age as the
driving parameter underlying the mass-step. If true, this would have a
significant impact for cosmology\nn{, not only because there would be a redshift
drift magnitude correction to apply, but also bevause it might strongly
vary} \yc{\textit{just the existence of a redshift-drift mag.\ correction is
of importance here, even if it does not vary}} \nn{\textit{now taken
into account}} \citep{rigault2013, childress2014, scolnic2018a}. Yet, the
importance of local SN environmental studies remains highly debated
\cite[e.g.][]{jones2015, jones2019} and especially the impact it
\yc{\textit{what does `it' refer to, the `importance'?}} \nn{\textit{no, to the
SN environment. Do you think it needs rephrasing?}} has on the derivation of
H$_0$ \citep{jones2015, riess2016, riess2018, rose2019}.

\yc{\textit{Shouldn't this paragraph, introducing the concept of old and young
populations and lsSFR, be moved \emph{\nn{here}}?}} \nn{\textit{I agree. Now
done.}}

\nn{The concept of the SNe~Ia age dichotomy arose with the study of the SN~Ia
    rate. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008} have
    shown that the relative SNe~Ia rate in galaxies could be explained if two
    populations existed, one young, following the host star formation activity,
    and one old following the host stellar mass (the so called ``prompt and
    delayed'' or ``A+B'' model). In \cite{rigault2018} we used the \nnn{LsSFR
    (}specific Star Formation Rate at the SN location\nnn{)} to classify which
    are the younger (those with a high LsSFR) and which are the older (those
    with low LsSFR). Furthermore, since the first SNe~Ia host analyses, the SN
    stretch has been known to be strongly correlated with the SN host properties
    \citep{hamuy1996, hamuy2000}, \nnn{correlation that} \nnn{\sout{which}}
    \yc{\textit{what does this ``which'' refer to?}} \nnn{\textit{is that
    better?}} has been extensively confirmed since \citep[e.g.][]{neill2009,
sullivan2010, lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013,
rigault2013, pan2014, kim19}. Following the ``A+B'' model and the connection
between SN stretch and host properties, \cite{howell2007} first discussed the
potential redshift drift of the SN stretch distribution. In this paper we
revisit this analysis with the most recent SNe~Ia dataset.}

In this paper, we take a step aside to probe the validity of our modeling of the
SN population, which we claim to be constituted of two age-populations
\citep{rigault2013,rigault2015,rigault2018}: one old and one younger, the former
having on average lower lightcurve stretches and being brighter after
standardization. We use the correlation between the SN age, as probed by local
specific star formation rate \citep{rigault2018}, and the SN stretch to model
the expected evolution of the underlying SN stretch distribution as a function
of redshift. This modeling relies on three assumptions: (1) there are two
distinct populations of SNe~Ia; (2) the relative fraction of each of these
populations as a function of redshift follows the model presented in
\cite{rigault2018} and (3) the underlying distribution of stretch for each age
sample \sout{\yc{is constant}} \nn{is a fixed model} and does not significantly
evolve with time. This paper aims at testing this specific model with datasets
from the literature.

We present in section~\ref{sec:sample} the sample we are using for this
analysis, derived from the Pantheon catalog \citep{scolnic2018a}. We discuss the
importance of obtaining a ``complete'' sample, i.e.\ representative of the true
underlying SNe Ia distribution, and how we build one from the Pantheon sample.
We then present in section~\ref{sec:modeling} our modeling of the distribution
of stretch\nn{, using \sout{as a function of}} redshift\nn{-}based models for
the stretch distribution of young and old SNe~Ia \yc{\textit{what??? rephrase}}
\nn{\textit{Is that better?}}. Our results are presented in
section~\ref{sec:results}, where we test whether the SN stretch distribution
evolves as a function of redshift and if the aforementioned age model is in good
agreement with this evolution. Consequences for cosmology of our results are
briefly discussed in section~\ref{sec:bbc} and we conclude in
section~\ref{sec:ccl}.

@@

We base our analysis on the most recent \yc{comprehensive}\nn{\textit{What does
it mean? Are there SNe~Ia compilation that are not comprehensive?}} SNe~Ia
compilation, the Pantheon catalog from \cite{scolnic2018a}. A naive approach to
test the SN stretch redshift drift would be to simply compare the observed SN
stretch distributions in a few bins of redshift. In practice, however,
\yc{differential} \nn{\textit{Do you mean ``different'', or that it is based on
        differences (the definition of ``differential'' as given by my search
engine)?}} selection effects are affecting the observed SN stretch
distributions. Indeed, because the observed SN~Ia magnitude correlates with the
lightcurve stretch (and color), the first SNe~Ia that a magnitude-limited survey
will miss are the lowest-stretch (and reddest) ones. Consequently, if
magnitude-related selection effects are not accounted for, one might confuse
true population drift with survey properties, and conversely.

@@

We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color of SNe~Ia
from the following surveys: PanStarrs \citep[PS1][]{rest2014}, the Sloan Digital
Sky Survey \citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plane with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates the full
distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011} and
\citet{campbell2013} for similar contours, the second using a more conservative
$|c| \leq 0.2$ cut.  Assuming the SN absolute magnitude with $x_1=0$ and $c=0$
is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$
along the aforementioned ellipse given the standardization coefficient
$\alpha=0.156$ and $\beta=3.14$ from \cite{scolnic2018a}: the faintest SN~Ia is
that with $(x_1=-1.65, c=0.25)$ and an absolute standardized magnitude at peak
in Bessel $B$ band of $M^{t_0}_{min} = -18.31$~mag. Since one ought to detect
this object typically \nn{\sout{a week} 5 days before and \sout{10 days} a week}
after peak to build a suitable lightcurve, the effective limiting standardized
absolute magnitude is approximately $M_{lim} = -18.00$~mag. Hence, given the
magnitude limit $m_{lim}$ of a magnitude limited survey, one can derive the
maximum redshift $z_{lim}$ above which the faintest SNe~Ia will be missed using
the relation between apparent magnitude, redshift and absolute magnitude
$\mu(z_{lim}) = m_{lim} - M_{lim}$.

% YC: Figs and Tables *after* their 1st ref. in the text!
\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
    \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$)
        lightcurve parameters of SNe~Ia from the SDSS, PS1 and SNLS samples of
        the Pantheon catalog. The individual SNe are shown as blue dots
        \nn{\sout{and a 2D histogram is shown in gray to highlight point
        density}}. \yc{\textit{I don't think it's really needed}} \nn{\textit{Ok,
changed. Better?}}. The ellipse $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by
the corresponding standardized absolute magnitude using the $\alpha$ and $\beta$
coefficients from \cite{scolnic2018a}. The grey diagonal lines represent the
$(x_1, c)$ evolution for $m = m_{lim}$, for $z$ between $0.50$ and $z=1.70$
using SNLS's $m_{lim}$ of $24.8$~mag.}
    \label{fig:maglim}
\end{figure}

@@

In a similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However, the
SDSS surveys were more sensitive to limited spectroscopic resources;
\cite{kessler2009} pointed out that during the first year of SDSS, SNe~Ia with
$r-mag<20.5$ were favored for spectroscopic follow-up, corresponding to a
redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were available, and \cite{kessler2009} and
\cite{dilday2008} show a relative completeness \yc{\textit{what is a
\emph{relative completeness}???}} \nn{\textit{In Dilday 2008, they give a median
redshift of <z> = .22 for spectroscopically confirmed SNe~Ia}} up to
$z_{lim}=0.2$. Following these analyses, we will use $z_{lim}=0.2$ as the
baseline SDSS redshift limit.

\nn{\sout{For the rest of the analysis, we will consider the aforementioned redshift
limits, as well as more conservative values for each of these surveys, namely
$z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for PS1 and $z_{lim}=0.55$ for SNLS.}}
\yc{why do you remind the conservative values, and not the fiducial ones? Just
refer to Table~1.}

The complete sample selection is summarized in Table~\ref{tab:sample}, and the
redshift distribution of these three surveys is shown in Fig.~\ref{fig:cuts}.
As expected, the selected redshift limits roughly correspond to the peak of
these histograms.

@@

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
and~4.2.2; \mr{SNe~Ia time series are published in \citealt{Saunders2020}}
\nn{should add Aldering in prep. and Pander in prep. as suggested by Greg}). The
spectroscopic follow-up was done over a redshift range of $0.02 < z <
0.09$~\citep[as in][]{rigault2018}, while the search was much deeper. As such,
these SNe~Ia are assumed to be a random sampling of the underlying SN
population. The SNfactory sample is particularly useful for studying SN property
drift, as it enables us to have a large SN~Ia sample at $z<0.1$.

@@

We present the stretch distribution and redshift histogram of these five surveys
up to their respective $z_{lim}$ in Fig.~\ref{fig:sample}. \nn{The combined
samples' mean stretches in bins of redshift is also presented
in Fig.~\ref{fig:modelall}, where we see that when considering
selection-bias-free SN~Ia samples, SNe~Ia at higher redshift have on average
larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower redshift
($-0.17\pm 0.10$ at $z\sim0.05$). That is the evolution that we attempt to model
in the next section.} \nnn{\textit{Last part added following Yannick's comment
about the fact that this result is model-independent; see last paragraph before
section 3.2}}

\yc{\textit{You should also present/refer to Fig.5, showing the significant
evolution of mean stretch as function of $z$ in complete samples. Next Sec. aims
at modeling this evolution.}} \nn{\textit{Done.}}

@@

\subsection{``Base'' underlying stretch distribution}\label{sec:basemodel}

\yc{\textit{There is a generic confusion between ``SN stretch'' (individual SN,
as in Eq.3/4) and something like ``SN stretch distribution'' (population, as in
Eq.2).  This is confusing...}}

To model the evolution of the \nn{global} SN stretch distribution as a function
of redshift, given our aforementioned model of the evolution of the fraction of
younger and older SNe~Ia with \yc{\sout{redshift} cosmic time} \nn{\textit{Is it
        really common to call $z$ the ``cosmic time''? What does this
modification bring to the understanding of this evolution? We might make the
confusion between the ``age'' of a SNe, and the ``time'' of its explosion given
by the redshift...}}, we need to model the SN stretch distribution for each age
subsample. \yc{\textit{That paragraph sounds like a tautology...}}
\nn{\textit{Added ``global'' so that we see its meaning}}

\yc{\textit{I suggest we remove the use of ``we'' when referring to previous papers,
like ``In R+18, we did...'' to be replaced by ``R+18 did...''}} \nn{\textit{That
has been a recurring question. This paper is part of a series of paper, so it
seemed ok, but for that I'll trust the common way, though I feel like lying by
making it sound impartial when we actually use ``our'' results. People should
know that we take our previous work as a baseline.}}

@@

When combined to the previously described prediction of the evolution of younger
SNe~Ia as a function of redshift (Eq.~\ref{eq:delta}), our model $X_1(z)$ for
the underlying stretch distribution as a function of redshift is written
\begin{multline}\label{eq:stretchz}
    X_1(z) = \delta(z) \times \mathcal{N}(\mu_1,\sigma_1{}^2) + {} \\
    (1-\delta(z)) \times \left[ a\times\mathcal{N}(\mu_1,\sigma_1{}^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2{}^2) \right].
\end{multline}

\yc{\textit{In fact, not clear what this eq. does here: it is supposed to be
superseded by Eq.3/4. Maybe it should be moved to Sec.3.1.2 when we use
$\delta(z)$ as a proxy to $y^i$.}} 
\nn{\textit{It should not! The evolution as a function of $z$ \underline{doesn't
depend on how we determine the parameters}. \textbf{This} is what people should
use for the ``Base'' model, and \textbf{then} there a two ways to get the
parameters. If you think there is a clearer way for people to understand this,
please tell us. All 3 eq. (2), (3) and (4) are of the same importance and are
used in the .py we wrote.}}

The maximum-likelihood estimate of the 5~free parameters
$\vec{\theta}\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained
by minimizing the following pseudo-$\chi^2$:
\begin{equation}\label{eq:likelihood}
    -2\ln(L) = -2 \sum_i \ln \prob{x^{i}_{1}}{\vec{\theta};
    \mathrm{d}x^{i}_{1}, y^i},
\end{equation}
with
\begin{align}\label{eq:likelihoodsnf}
    \prob{x^i_1}{\vec{\theta}; \mathrm{d}x^i_1, y^i} =
    y^i & \times
    \mathcal{N}\left(x^i_1 \mid \mu_1, \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    (1-y^i) &\times \bigg[
    a \times \mathcal{N}\left(x^i_1 \mid \mu_1,
    \sigma_1{}^2+\mathrm{d}x^i_1{}^2\right) +
    \nonumber\\
    & (1-a) \times \mathcal{N}\left(x^i_1 \mid \mu_2,
    \sigma_2{}^{2}+\mathrm{d}x^i_1{}^2\right) \bigg],
\end{align}
